### 2020-11-17 paper.tex
title change + NU sources

\documentclass[12pt]{article}
\title{Suffering Through Sanctions: An Application of Utilitarianism to Coercive Economic Sanctions}
\author{Word Count: 100}
\date{\today}

\usepackage{setspace}
\usepackage[utf8]{inputenc}
\usepackage{CormorantGaramond}
\usepackage{hyperref}
\usepackage{indentfirst}
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage{microtype}
\usepackage{blindtext}
\usepackage[margin=1in]{geometry}
\usepackage[american]{babel}
\usepackage{csquotes}
\usepackage[style=mla]{biblatex}
\addbibresource{library.bib}
\pagestyle{fancy}
\fancyhf{}
\rhead{Suffering Through Sanctions \thepage}
\setlength{\headheight}{15pt}
\linespread{2}
\raggedright{}
\setlength\parindent{0.5in}

\begin{document}
\maketitle

\Blindtext{}

% \printbibliography{}
\end{document}